Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Patterns in Australian Spatial Statistics

### 2025-07-09 paper/kobakian-cook-anzjs.qmd
done with all - needs a careful read and references cleaned up

<!-- Other cartograms -->
Alternative algorithms make various trade offs between familiar shapes and representation of geographic units. The non-contiguous cartogram method [@NAC] keeps the shapes of geographic units intact, and changes the size of the shape. This method disconnects areas creating empty space on the display losing the continuity of the spatial display of the statistic. The Dorling cartogram [@ACTUC] represents each unit as a circle, sized according to the value of the statistic. The neighbour relationships are mostly maintained by how the circles touch. A similar approach was pioneered by @RSCW, using rectangles that tile to align borders of neighbours [@CDWCS]. There have been thorough reviews of the array of methods, as suitable for cancer atlas displays [@review;@BCM].

@@

 - **Plot choice**: the number of the plot that they deemed to be most different from the others.
 - **Reason**: one of "Clusters of colour", "Colour trend across the areas", "Big differences between neighbouring areas", "All areas have similar colours" or "None of these reasons". 
 - **Certainty**: how certain that their choice is different from the others, on a scale of 1-5.

@@

Participants were asked why they had made their plot choice and were able to select from a set of suggested reasons. 
"Colour trend across the areas" was the most common selection for NW-SE trend displays (@tbl-reason).

@@

The choice "None of these reasons" was used as the default value to minimise noise from participants who did not select a response.

```{r}
#| label: tbl-reason
#| tbl-cap: "Summary of the reasons that participants wrote for their choice of plot when looking at each trend model shown in choropleth and hexagon tile maps."
# Qualitative analysis of reason
reasons_smry <- d %>% 
  mutate(reason = ifelse(reason =="0.0", "no reason", reason)) %>% 
  mutate(Detected = ifelse(detect_f == "Detected? Yes", "Yes", "No"),
    Trend = trend) %>% 
  group_by(Trend, Detected, type) %>% 
  count(reason) %>%
  mutate(prop = round(n/sum(n), 2)) %>% 
  pivot_wider(names_from = reason, 
              values_from = prop, 
              values_fill = 0)

#  mutate(prop = round(n/sum(n), 2), r_prop = paste0(reason, ":", prop)) %>% 
#  top_n(1, n) %>% summarise(reasons = paste(reason, collapse=", ")) %>% 
#  pivot_wider(names_from = c("type"), values_from = c("reasons")) %>% 
#  ungroup() %>% 
reasons_smry |>  
  knitr::kable(format = "latex", booktabs = TRUE,
    linesep = c("", "\\addlinespace", "", "\\addlinespace","")) %>% 
  #kable_styling(bootstrap_options = "hold_position") %>% 
  collapse_rows(., columns = 1)
```

@@

The intention of this study was to contrast the use of the choropleth map and the hexagon tile map.
The visual inference lineup protocol was employed to contrast the effectiveness of the displays.
The results have shown that overall the use of the hexagon tile map display allows participants to find the data plot in the lineup more often. Using the visual inference protocol this result can be extended to show that it is a valid alternative display to communicate spatial distributions of population related data.

@@

The increasing population densities of capital cities despite large land area exacerbates the difference in the smallest and largest communities.
The population density structure of Australia can be considered similar to that of Canada, New Zealand and many other countries. Therefore, this display is not only relevant to Australia, but all nations or population distributions that experience densely populated cities separated by vast rural expanses.

- Add Election results data + plus other statistics, 
- new and better hexagon layout software 
- Other plot/map type comparisons, insets, cartograms, statistical plots along with maps, especially interactive
- eye-tracking in association with lineups

>>>>

<!-- Other cartograms -->
Alternative algorithms make various trade offs between familiar shapes and representation of geographic units. The non-contiguous cartogram method [@NAC] keeps the shapes of geographic units intact, and changes the size of the shape. This method disconnects areas creating empty space on the display losing the continuity of the spatial display of the statistic. The Dorling cartogram [@ACTUC] represents each unit as a circle, sized according to the value of the statistic. The neighbour relationships are mostly maintained by how the circles touch. A similar approach was pioneered by @RSCW, using rectangles that tile to align borders of neighbours [@CDWCS]. There have been thorough reviews of the array of methods, as suitable for cancer atlas displays [e.g. @KCR;@BCM], and experiments demonstrating cartograms to be more effective than choropleth maps [@KFIF].

@@

 - **Plot choice**: the number of the plot that they deemed to be most different from the others.
 - **Reason**: one of "Clusters of colour", "Colour trend across the areas", "Big differences between neighbouring areas", "All areas have similar colours" or "None of these reasons". Note providing restricted list of reasons rather than free text encourages a response because it is easy. The list needs to contain the primary expected reasons and other potential reasons need to be added so that it does not bias the participants' behaviour.
 - **Certainty**: how certain that their choice is different from the others, on a scale of 1-5.

@@

@tbl-reason summarises the reasons that participants gave for their choices: "clusters" = "Clusters of colour", "trend" = "Colour trend across the areas", "consistent" = "All areas have similar colours", "hotpots" = "Big differences between neighbouring areas", "none" = "None of these reasons". These proportions are computed separately for each trend, map type and whether the participant detected the data plot. With All Cities and Three Cities, when correct, participants tended to select 'consistent' with the choropleth, and 'clusters' with the hexagon tile map. With the NW-SE trend, 'trend' was primarily selected for the choropleth, but 'clusters' was the primary reason for the hexgon tile map. The primary reasons are similar when the participant did not detect the data plot.

The results when the data plot was detected are as expected, but that they are similar to the not detected group is interesting. It suggests that with the choropleth people are trying to read contiguous patterns, while the hexagon tile map is being read for pockets of differences. This may be due the hexagon tile map being non-contiguous.

<!--Participants were asked why they had made their plot choice and were able to select from a set of suggested reasons. 
"Colour trend across the areas" was the most common selection for NW-SE trend displays (@tbl-reason).

@@

The choice "None of these reasons" was used as the default value to minimise noise from participants who did not select a response. -->

```{r}
#| label: tbl-reason
#| tbl-cap: "Proportion of participants chose each reason for their choice of plot, computed separately for each Trend and Map Type level, and whether the data plot was detected. The primary reason when participants were evaluating the choropleth map was 'consistent' or 'trend', but for the hexagon tile map it was 'clusters', when they detected the data plot."
# Qualitative analysis of reason
reasons_smry <- d %>% 
  mutate(reason = ifelse(reason =="0.0", "no reason", reason)) %>% 
  mutate(Detected = ifelse(detect_f == "Detected? Yes", "Yes", "No"),
    Trend = trend) %>% 
  group_by(Trend, Detected, type) %>% 
  count(reason) %>%
  ungroup() %>%
  group_by(Trend, Detected, type) %>%
  mutate(prop = n/sum(n)) %>% 
  ungroup() %>%
  select(Trend, Detected, type, reason, prop) %>%
  mutate(prop = round(prop, 2)) %>%
  pivot_wider(names_from = reason, 
              values_from = prop, 
              values_fill = 0) %>%
  select(Trend, Detected, type, clusters, trend, consistent, hotspots, `no reason`)

#  mutate(prop = round(n/sum(n), 2), r_prop = paste0(reason, ":", prop)) %>% 
#  top_n(1, n) %>% summarise(reasons = paste(reason, collapse=", ")) %>% 
#  pivot_wider(names_from = c("type"), values_from = c("reasons")) %>% 
#  ungroup() %>% 
reasons_smry |>  
  arrange(desc(Detected), Trend, type) |>
  knitr::kable(format = "latex", booktabs = TRUE,
               col.names = c("Trend", "Detect", "Type", 
                             "clusters", "trend", "consistent",
                             "hotspots", "none"),
                 align = "lllrrrrr",
    linesep = c("", "\\addlinespace", "", "\\addlinespace","", "\\addlinespace","", "\\addlinespace","", "\\addlinespace","")) #%>% 
  #kable_styling(bootstrap_options = "hold_position") %>% 
  #collapse_rows(., columns = 1)
```

@@

This study provides evidence that the hexagon tile map is superior to a choropleth map for communicating population statistics, for Australia. While the cartogram has been established as better than a choropleth map, cartograms do not work for the vast disparity between population density and geographic area in Australia. The hexagon tile map was developed to provide a possible solution, and this study demonstrates that it has potential.

The R package `sugarbag` can be used to generate a hexagon tile map. It can be used for any spatial polygon data, so is applicable to other countries or geographic areas.

One of the strengths but weaknesses of the hexagon tile map is that it is non-contiguous. Large rural areas are represented by isolated hexagons. This is why we expected that the hexagon tile map might not work well to detect large-scale spatial trend ("NW-SE"). The primary reason for producing the non-contiguous display was to preserve geography sufficiently for the reader to easily recognise the location. This is a strength, and allows a map of the country to be drawn underneath the hexagons. It appears to not inhibit reading of the spatial distribution based on this experiment. However, there is considerable room for improving the algorithm and exploring some variations. These might include increasing the size of isolated hexagons, or collecting multiple hexagons together.

The manner in which hexagons are exploded out from the city centres is another direction of research. Ideally, the location of the hexagons should be close to their original location but this is hard to control and measure. The current algorithm works sequentially to place hexagons, radially from a provided centre. There are likely better optimisation procedures that could impve the layout. For reading the spatial distribution, this is less important, but if the hexagon tile map is provided to users as an interactive tool, they will want to locate themselves in the plot. If the hexagon is not close to the true location it could be disconcerting.

Hexagon displays are growing in popularity. Two media outlets used variations of the hexagon displays to communicate the 2025 Australian federal election results [@abc-election2025;guardian-election2025]. Both are effective, but have inadequacies. The Guardian preserves geography and allows inner city results to be seen but the overall sense of the result is skewed because the large rural areas dominate the display. The ABC's contiguous hexagon tile map gives the correct sense of the final results but loses the shape of Australia, and some hexagons are far from their true location.

This experiment focused on comparing a new display, the hexagon tile map, against the standard display, choropleth. There are other options that could have been included in the study, such as the use of insets of dense population areas along with the chorpleth map, or the use of interactive graphics linking statistical charts with the choropleth. Keeping the scope of the study small was important to understand whether it was reasonable to recommend use of the hexagon tile map. Although we only tested on the Australian geography, the results should hold for other regions that have similarly disparities between population size and geographic size.

We would recommend doing follow-up studies that allow deeper understanding of how the different displays are read. For example, to understand the differences in how people read the choropleth map and the hexagon tile map as indicated by the different reasons given, could be done using an eyetracking experiment. @ZCHMR is an example of such an experiment where the manner in which people read lineups was examined.

<!--The visual inference lineup protocol was employed to contrast the effectiveness of the displays.
The results have shown that overall the use of the hexagon tile map display allows participants to find the data plot in the lineup more often. Using the visual inference protocol this result can be extended to show that it is a valid alternative display to communicate spatial distributions of population related data.

@@

The increasing population densities of capital cities despite large land area exacerbates the difference in the smallest and largest communities.
The population density structure of Australia can be considered similar to that of Canada, New Zealand and many other countries. Therefore, this display is not only relevant to Australia, but all nations or population distributions that experience densely populated cities separated by vast rural expanses.
-->

### 2025-07-09 paper/kobakian-cook-anzjs.tex
done with all - needs a careful read and references cleaned up

Alternative algorithms make various trade offs between familiar shapes
and representation of geographic units. The non-contiguous cartogram
method \citep{NAC} keeps the shapes of geographic units intact, and
changes the size of the shape. This method disconnects areas creating
empty space on the display losing the continuity of the spatial display
of the statistic. The Dorling cartogram \citep{ACTUC} represents each
unit as a circle, sized according to the value of the statistic. The
neighbour relationships are mostly maintained by how the circles touch.
A similar approach was pioneered by \citet{RSCW}, using rectangles that
tile to align borders of neighbours \citep{CDWCS}. There have been
thorough reviews of the array of methods, as suitable for cancer atlas
displays \citep{review, BCM}.

@@

\begin{itemize}
\tightlist
\item
  \textbf{Plot choice}: the number of the plot that they deemed to be
  most different from the others.
\item
  \textbf{Reason}: one of ``Clusters of colour'', ``Colour trend across
  the areas'', ``Big differences between neighbouring areas'', ``All
  areas have similar colours'' or ``None of these reasons''.
\item
  \textbf{Certainty}: how certain that their choice is different from
  the others, on a scale of 1-5.
\end{itemize}

@@

Participants were asked why they had made their plot choice and were
able to select from a set of suggested reasons. ``Colour trend across
the areas'' was the most common selection for NW-SE trend displays
(Table~\ref{tbl-reason}).

The reasons chosen by participants from the list provided to them varied
more when viewing choropleth displays than the hexagon tile map. The
hexagon tile map displays resulted in ``Clusters of colour'' as the most
common choice made by participants.

The choice ``None of these reasons'' was used as the default value to
minimise noise from participants who did not select a response.

@@

\caption{\label{tbl-reason}Summary of the reasons that participants
wrote for their choice of plot when looking at each trend model shown in
choropleth and hexagon tile maps.}

@@

\begin{tabular}{lllrrrrr}
\toprule
Trend & Detected & type & clusters & consistent & hotspots & no reason & trend\\
\midrule
 & No & Choro. & 50 & 12 & 11 & 9 & 56\\
\cmidrule{2-8}
 & No & Hex. & 27 & 6 & 5 & 15 & 22\\
\cmidrule{2-8}
 & Yes & Choro. & 14 & 14 & 3 & 4 & 7\\
\cmidrule{2-8}
\multirow{-4}{*}{\raggedright\arraybackslash All Cities} & Yes & Hex. & 44 & 4 & 14 & 8 & 35\\
\cmidrule{1-8}
 & No & Choro. & 54 & 12 & 17 & 16 & 74\\
\cmidrule{2-8}
 & No & Hex. & 42 & 6 & 11 & 15 & 34\\
\cmidrule{2-8}
 & Yes & Choro. & 1 & 4 & 0 & 0 & 2\\
\cmidrule{2-8}
\multirow{-4}{*}{\raggedright\arraybackslash Three Cities} & Yes & Hex. & 35 & 0 & 10 & 4 & 23\\
\cmidrule{1-8}
 & No & Choro. & 20 & 14 & 11 & 8 & 34\\
\cmidrule{2-8}
 & No & Hex. & 27 & 3 & 5 & 15 & 18\\
\cmidrule{2-8}
 & Yes & Choro. & 32 & 2 & 15 & 7 & 37\\
\cmidrule{2-8}
\multirow{-4}{*}{\raggedright\arraybackslash NW-SE} & Yes & Hex. & 56 & 12 & 6 & 6 & 32\\
\bottomrule
\end{tabular}

@@

The intention of this study was to contrast the use of the choropleth
map and the hexagon tile map. The visual inference lineup protocol was
employed to contrast the effectiveness of the displays. The results have
shown that overall the use of the hexagon tile map display allows
participants to find the data plot in the lineup more often. Using the
visual inference protocol this result can be extended to show that it is
a valid alternative display to communicate spatial distributions of
population related data.

We expected that the choropleth map would be superior for communicating
the spatial pattern of geographic distributions. The data suggest that
the participants perform slightly better or equally as well for each
replicate in each trend model across the two displays. Table II shows
that the difference in the mean detection rate for the two trend models
was 0.10.

The differences seen in the Figure~\ref{fig-detect-compare} plot.
Surprisingly the difference between map types for the geographic
distribution was significant at the 0.05 level, even though the effect
difference is small. It also showed that the hexagon tile map display
performs marginally better than the choropleth for all trend models.

While the significance of the difference in detection was the key focus
of this experiment, the secondary focus was the time taken by
participants. it was expected that the participants may take longer to
consider the hexagon tile map distribution but would be able to detect
the data plot in the lineup. The bimodal distributions seen in
Figure~\ref{fig-beeswarm} showed very little difference in the mean
evaluation times. As the maximum time of all of the distributions
approached 60 seconds it cannot be said that the participants' took
longer to evaluate the hexagon tile map displays.

The responses to the questions asked of participants included the reason
for their choice and the certainty around their choice.
Table~\ref{tbl-reason} shows high levels of certainty of 4 and 5 were
chosen by participants when looking at the population distributions in a
choropleth map display show that they were over confident when
attempting to find the real data plot in the choropleth map displays.
Participants performed better on the NW-SE distribution shown in the
choropleth display and were reasonably confident about their decisions.
The high levels of the mid range value of 3 could indicate that the
participant did not want to provide a response, as this was the default
value. Those who chose level 4 or 5 were equally likely to be correct
for the three cities lineups, but more likely to be correct than
incorrect for the other two trend models.

The colour scaling applied in Three cities and All cities displays
resulted in the rural areas of the real data plot appearing more blue or
yellow than the other plots in the lineups. Due to the consistent
colouring of rural areas in a choropleth display, the choice ``All areas
have similar colours'' was most common reason for a participants choice.
The All Cities displays coloured the inner-city areas of all capital
cities more red, this was observable to participants and explains the
equal choice of the city clusters or rural colour consistency. Choosing
``Clusters of colour'' was expected when participants viewed the Hexagon
tile map display of the All Cities and Three Cities distributions. It
was unexpected that it was also the most common reason for the NW-SE
hexagon tile map displays. Due to the spatial covariance introduced in
the smoothing, groups of similarly coloured hexagons were present in all
of the hexagon tile map displays. All Cities and Three Cities
distributions of real data trends had distinctly different patterns or
red inner-city areas, while some of the plots in each lineup may have
shared similar features.

\section{Conclusion}\label{conclusion}

The choropleth map display and the tessellated hexagon tile map have
been contrasted using the lineup protocol. The hexagon tile map was
significantly more effective for spotting a real population related data
trend model hidden in a lineup.

The hexagon tile map display should be considered as an alternative
visualisation method when communicating distributions that relate to the
population across a set of geographic units. As an additional display to
the familiar choropleth map, cancer atlas products may benefit from the
opportunity to allow exploration via an alternative display. The spatial
distributions used to test these displays were inspired by the real
spatially smoothed estimates of the cancer burden on Australian
communities. However, this technique may be extended to other population
related distributions, such as other diseases, or election results or
socioeconomic indicators.

The increasing population densities of capital cities despite large land
area exacerbates the difference in the smallest and largest communities.
The population density structure of Australia can be considered similar
to that of Canada, New Zealand and many other countries. Therefore, this
display is not only relevant to Australia, but all nations or population
distributions that experience densely populated cities separated by vast
rural expanses.

\begin{itemize}
\tightlist
\item
  Add Election results data + plus other statistics,
\item
  new and better hexagon layout software
\item
  Other plot/map type comparisons, insets, cartograms, statistical plots
  along with maps, especially interactive
\item
  eye-tracking in association with lineups
\end{itemize}

>>>>

Alternative algorithms make various trade offs between familiar shapes
and representation of geographic units. The non-contiguous cartogram
method \citep{NAC} keeps the shapes of geographic units intact, and
changes the size of the shape. This method disconnects areas creating
empty space on the display losing the continuity of the spatial display
of the statistic. The Dorling cartogram \citep{ACTUC} represents each
unit as a circle, sized according to the value of the statistic. The
neighbour relationships are mostly maintained by how the circles touch.
A similar approach was pioneered by \citet{RSCW}, using rectangles that
tile to align borders of neighbours \citep{CDWCS}. There have been
thorough reviews of the array of methods, as suitable for cancer atlas
displays \citep[e.g.][]{KCR, BCM}, and experiments demonstrating
cartograms to be more effective than choropleth maps \citep{KFIF}.

@@

\begin{itemize}
\tightlist
\item
  \textbf{Plot choice}: the number of the plot that they deemed to be
  most different from the others.
\item
  \textbf{Reason}: one of ``Clusters of colour'', ``Colour trend across
  the areas'', ``Big differences between neighbouring areas'', ``All
  areas have similar colours'' or ``None of these reasons''. Note
  providing restricted list of reasons rather than free text encourages
  a response because it is easy. The list needs to contain the primary
  expected reasons and other potential reasons need to be added so that
  it does not bias the participants' behaviour.
\item
  \textbf{Certainty}: how certain that their choice is different from
  the others, on a scale of 1-5.
\end{itemize}

@@

Table~\ref{tbl-reason} summarises the reasons that participants gave for
their choices: ``clusters'' = ``Clusters of colour'', ``trend'' =
``Colour trend across the areas'', ``consistent'' = ``All areas have
similar colours'', ``hotpots'' = ``Big differences between neighbouring
areas'', ``none'' = ``None of these reasons''. These proportions are
computed separately for each trend, map type and whether the participant
detected the data plot. With All Cities and Three Cities, when correct,
participants tended to select `consistent' with the choropleth, and
`clusters' with the hexagon tile map. With the NW-SE trend, `trend' was
primarily selected for the choropleth, but `clusters' was the primary
reason for the hexgon tile map. The primary reasons are similar when the
participant did not detect the data plot.

The results when the data plot was detected are as expected, but that
they are similar to the not detected group is interesting. It suggests
that with the choropleth people are trying to read contiguous patterns,
while the hexagon tile map is being read for pockets of differences.
This may be due the hexagon tile map being non-contiguous.

@@

\caption{\label{tbl-reason}Proportion of participants chose each reason
for their choice of plot, computed separately for each Trend and Map
Type level, and whether the data plot was detected. The primary reason
when participants were evaluating the choropleth map was `consistent' or
`trend', but for the hexagon tile map it was `clusters', when they
detected the data plot.}

@@

\begin{tabular}{lllrrrrr}
\toprule
Trend & Detect & Type & clusters & trend & consistent & hotspots & none\\
\midrule
All Cities & Yes & Choro. & 0.33 & 0.17 & 0.33 & 0.07 & 0.10\\
All Cities & Yes & Hex. & 0.42 & 0.33 & 0.04 & 0.13 & 0.08\\
\addlinespace
Three Cities & Yes & Choro. & 0.14 & 0.29 & 0.57 & 0.00 & 0.00\\
Three Cities & Yes & Hex. & 0.49 & 0.32 & 0.00 & 0.14 & 0.06\\
\addlinespace
NW-SE & Yes & Choro. & 0.34 & 0.40 & 0.02 & 0.16 & 0.08\\
NW-SE & Yes & Hex. & 0.50 & 0.29 & 0.11 & 0.05 & 0.05\\
\addlinespace
All Cities & No & Choro. & 0.36 & 0.41 & 0.09 & 0.08 & 0.07\\
All Cities & No & Hex. & 0.36 & 0.29 & 0.08 & 0.07 & 0.20\\
\addlinespace
Three Cities & No & Choro. & 0.31 & 0.43 & 0.07 & 0.10 & 0.09\\
Three Cities & No & Hex. & 0.39 & 0.31 & 0.06 & 0.10 & 0.14\\
\addlinespace
NW-SE & No & Choro. & 0.23 & 0.39 & 0.16 & 0.13 & 0.09\\
NW-SE & No & Hex. & 0.40 & 0.26 & 0.04 & 0.07 & 0.22\\
\bottomrule
\end{tabular}

@@

This study provides evidence that the hexagon tile map is superior to a
choropleth map for communicating population statistics, for Australia.
While the cartogram has been established as better than a choropleth
map, cartograms do not work for the vast disparity between population
density and geographic area in Australia. The hexagon tile map was
developed to provide a possible solution, and this study demonstrates
that it has potential.

The R package \texttt{sugarbag} can be used to generate a hexagon tile
map. It can be used for any spatial polygon data, so is applicable to
other countries or geographic areas.

One of the strengths but weaknesses of the hexagon tile map is that it
is non-contiguous. Large rural areas are represented by isolated
hexagons. This is why we expected that the hexagon tile map might not
work well to detect large-scale spatial trend (``NW-SE''). The primary
reason for producing the non-contiguous display was to preserve
geography sufficiently for the reader to easily recognise the location.
This is a strength, and allows a map of the country to be drawn
underneath the hexagons. It appears to not inhibit reading of the
spatial distribution based on this experiment. However, there is
considerable room for improving the algorithm and exploring some
variations. These might include increasing the size of isolated
hexagons, or collecting multiple hexagons together.

The manner in which hexagons are exploded out from the city centres is
another direction of research. Ideally, the location of the hexagons
should be close to their original location but this is hard to control
and measure. The current algorithm works sequentially to place hexagons,
radially from a provided centre. There are likely better optimisation
procedures that could impve the layout. For reading the spatial
distribution, this is less important, but if the hexagon tile map is
provided to users as an interactive tool, they will want to locate
themselves in the plot. If the hexagon is not close to the true location
it could be disconcerting.

Hexagon displays are growing in popularity. Two media outlets used
variations of the hexagon displays to communicate the 2025 Australian
federal election results
{[}\citet{abc-election2025};guardian-election2025{]}. Both are
effective, but have inadequacies. The Guardian preserves geography and
allows inner city results to be seen but the overall sense of the result
is skewed because the large rural areas dominate the display. The ABC's
contiguous hexagon tile map gives the correct sense of the final results
but loses the shape of Australia, and some hexagons are far from their
true location.

This experiment focused on comparing a new display, the hexagon tile
map, against the standard display, choropleth. There are other options
that could have been included in the study, such as the use of insets of
dense population areas along with the chorpleth map, or the use of
interactive graphics linking statistical charts with the choropleth.
Keeping the scope of the study small was important to understand whether
it was reasonable to recommend use of the hexagon tile map. Although we
only tested on the Australian geography, the results should hold for
other regions that have similarly disparities between population size
and geographic size.

We would recommend doing follow-up studies that allow deeper
understanding of how the different displays are read. For example, to
understand the differences in how people read the choropleth map and the
hexagon tile map as indicated by the different reasons given, could be
done using an eyetracking experiment. \citet{ZCHMR} is an example of
such an experiment where the manner in which people read lineups was
examined.
